---
title: " synergy"
output: html_notebook
---
###TODO 
* why the same function fails at ZIP. It doesn't? i dont get it
* melting(?) output into long. Call python module into R. Eaasy
* add cell_line info

load libs
```{r}
library(synergyfinder)
library(naniar) # probably not needed
library(purrr) # probably not needed
library(nleqslv) # for debug the drm
```
basic read input using Alina's sorted data with celllines added (to help differentiate between batch1 and batch3)

```{r}
read.csv('/Users/zagidull/Desktop/processed_sorted_with_cellline.csv',sep = ";",
         colClasses = c('numeric', 'numeric','character', 'character','numeric','character','numeric','numeric',
                        'character','character','numeric','numeric')) -> input.sample

reshaped <- ReshapeData(input.sample,data.type = 'inhibition')
```
I know that 104 combo throws error when calculating Loewe synergy score, because i do 
```{r}
debugonce(Loewe)
options(error = recover)
dump.frames(dumpto = "last.dump", to.file = FALSE,include.GlobalEnv = FALSE)
debugger(dump = last.dump)
```
here i compose a new variable that consist of dose.response.mat and drug.pairs that throw in the error
```{r}
reshaped$dose.response.mats[104] -> oshibka
reshaped$drug.pairs[104,] -> imena
erl <- list()
erl$dose.response.mats <- oshibka
erl$drug.pairs <- imena
```
let's create a test case that we know works. It is for drug combo 103
```{r}
reshaped$dose.response.mats[103] -> neoshibka
reshaped$drug.pairs[103,] -> neimena
neerl <- list()
neerl$dose.response.mats <- neoshibka
neerl$drug.pairs <- neimena
```
let's create a case when both combos work
```{r}
reshaped$dose.response.mats[102] -> eneoshibka
reshaped$drug.pairs[102,] -> eneimena
eneerl <- list()
eneerl$dose.response.mats <- eneoshibka
eneerl$drug.pairs <- eneimena
enew <- list()
enew <- eneerl
enew$drug.pairs[2,] <- neerl$drug.pairs[1,]
enew$dose.response.mats[2] <- neerl$dose.response.mats[1]
```
let's create a case when out of two combos only one works
```{r}
new <- list()
new <- neerl
new$drug.pairs[2,] <- erl$drug.pairs[1,]
new$dose.response.mats[2] <- erl$dose.response.mats[1]
```

to test it throws the error I do
```{r}
today.Loewe <- CalculateSynergy(data = erl,method = "Loewe",correction = TRUE, Emin = 0, Emax = 100)
```
this throws error 
Error in nleqslv(max(drug.col.par[2] + 1, drug.row.par[2] + 1), eq, method = "Newton") : 
  initial value of fn function contains non-finite values (starting at index=1)
  Check initial x and/or correctness of function
Let's try to wrap it in tryCatch() to make it return NA
```{r}
irr2 <- function (x) {
  out <- tryCatch(CalculateSynergy(data = x,method = "Loewe",correction = TRUE, Emin = 0, Emax = 100), error = function(e) NA)
  return(out)
}
```
this works creating a dummy empty NA matrix for when there is one pair of drugs
```{r}
irr3 <- function (x) {
# scores <- list()
  method <- "Loewe"
 
  out <- tryCatch(CalculateSynergy(data = x,method = "Loewe",correction = TRUE, Emin = 0, Emax = 100), error = function(e) {
  x$method <- method
  x$scores <- x$dose.response.mats
#  x$scores[!is.na(x$scores)] <- NA
#  x$scores <- replace(x$scores, !is.na(x$scores), NA)
#  purrr::map(!is.na(x$scores), NA)
  x$scores <- lapply(x$scores, function(z) ifelse(!is.na(z), NA, z))
  return(x)
  })
  return(out)
}
```
here is the original of Loewe function
```{r}
Loewe_modded <- function (response.mat, correction = TRUE, Emin = NA, Emax = NA, 
    nan.handle = c("LL4", "L4")) 
{
    if (correction) {
        response.mat <- BaselineCorrectionSD(response.mat, Emin = Emin, 
            Emax = Emax, nan.handle)$corrected.mat
    }
    single.fit <- FittingSingleDrug(response.mat)
    drug.col.model <- single.fit$drug.col.model
    drug.col.par <- coef(drug.col.model)
    d1.fun <- function(conc, drug.col.par) {
        (drug.col.par[3] + drug.col.par[2] * (conc/drug.col.par[4])^drug.col.par[1])/(1 + 
            (conc/drug.col.par[4])^drug.col.par[1])
    }
    drug.row.model <- single.fit$drug.row.model
    drug.row.par <- coef(drug.row.model)
    d2.fun <- function(conc, drug.row.par) {
        (drug.row.par[3] + drug.row.par[2] * (conc/drug.row.par[4])^drug.row.par[1])/(1 + 
            (conc/drug.row.par[4])^drug.row.par[1])
    }
    row.conc <- as.numeric(rownames(response.mat))[-1]
    col.conc <- as.numeric(colnames(response.mat))[-1]
    loewe.mat <- response.mat
    for (i in 1:length(col.conc)) {
        for (j in 1:length(row.conc)) {
            x1 <- col.conc[i]
            x2 <- row.conc[j]
            eq <- function(x) {
                x1/(drug.col.par[4] * (((x - drug.col.par[3])/(drug.col.par[2] - 
                  x))^(1/drug.col.par[1]))) + x2/(drug.row.par[4] * 
                  (((x - drug.row.par[3])/(drug.row.par[2] - 
                    x))^(1/drug.row.par[1]))) - 1
            }
            slv <- nleqslv(max(drug.col.par[2] + 1, drug.row.par[2] + 
                1), eq, method = "Newton")
            if (slv$termcd == 1) {
                loewe.mat[j + 1, i + 1] <- slv$x
            }
            else {
                y.loewe1 <- d1.fun(x1 + x2, drug.col.par)
                y.loewe2 <- d2.fun(x1 + x2, drug.row.par)
                loewe.mat[j + 1, i + 1] <- ifelse(y.loewe1 > 
                  y.loewe2, y.loewe1, y.loewe2)
            }
        }
    }
    return(response.mat - loewe.mat)
}
```
here is Loewe is modded version.
```{r}
Loewe <- function (response.mat,correction, Emin = NA, Emax = NA,nan.handle = c("LL4", "L4")) {
# scores <- list()
#  method <- "Loewe"
  out <- tryCatch(Loewe_modded(response.mat,correction = correction, Emin = Emin, Emax = Emax, 
    nan.handle = nan.handle), error = function(e) NA)
#  {
#  x$method <- method
#  x$scores <- x$dose.response.mats
#  x$scores[!is.na(x$scores)] <- NA
#  x$scores <- replace(x$scores, !is.na(x$scores), NA)
#  purrr::map(!is.na(x$scores), NA)
#  x$scores <- lapply(x$scores, function(z) ifelse(!is.na(z), NA, z))
#  return(x)
#  })
  return(out)
}
```
here is calculate synergy with added tryCatch to Loewe method
```{r}
CalculateSynergy <- function (data, method = "ZIP", correction = TRUE, Emin = 0, 
    Emax = 100, nan.handle = c("LL4", "L4")) 
{
    if (!is.list(data)) {
        stop("Input data is not a list format!")
    }
    if (!method %in% c("ZIP", "HSA", "Bliss", "Loewe")) {
        stop("The method parameter can only be one of the following: ZIP, HSA, Bliss and Loewe.")
    }
    dose.response.mats <- data$dose.response.mats
    num.pairs <- length(dose.response.mats)
    scores <- list()
    nan.handle <- match.arg(nan.handle)
    for (i in 1:num.pairs) {
        response.mat <- dose.response.mats[[i]]
        scores[[i]] <- switch(method, 
            ZIP = ZIP(response.mat = response.mat, correction = correction, Emin = Emin, Emax = Emax), 
            HSA = HSA(response.mat, correction, Emin = Emin, Emax = Emax, nan.handle), 
            Bliss = Bliss(response.mat, correction, Emin = Emin, Emax = Emax, nan.handle), 
            Loewe = Loewe(response.mat, correction, Emin = Emin, Emax = Emax, nan.handle))
    }
    data$scores <- scores
    data$method <- method
    return(data)
}

```
modded ZIP. 
```{r}
ZIP_modded <- function (response.mat, correction = TRUE, Emin = 0, Emax = 100, 
    nan.handle = c("LL4", "L4")) 
{
    if (correction) {
        nan.handle <- match.arg(nan.handle)
        response.mat <- BaselineCorrectionSD(response.mat, NA, 
            NA, nan.handle)$corrected.mat
    }
    single.fitted <- FittingSingleDrug(response.mat, fixed = c(NA, 
        Emin, Emax, NA), nan.handle)
    drug.col.response <- single.fitted$drug.col.fitted
    drug.row.response <- single.fitted$drug.row.fitted
    updated.single.mat <- mat.or.vec(nrow(response.mat), ncol(response.mat))
    colnames(updated.single.mat) <- colnames(response.mat)
    rownames(updated.single.mat) <- rownames(response.mat)
    updated.single.mat[1, c(2:ncol(response.mat))] <- drug.col.response
    updated.single.mat[c(2:nrow(response.mat)), 1] <- drug.row.response
    updated.col.mat <- updated.single.mat
    for (i in 2:ncol(response.mat)) {
        tmp <- as.data.frame(mat.or.vec(nrow(response.mat) - 
            1, 0))
        tmp$dose <- as.numeric(rownames(response.mat)[-1])
        tmp$inhibition <- response.mat[c(2:nrow(response.mat)), 
            i]
        tmp.min <- updated.single.mat[1, i]
        if (var(tmp$inhibition, na.rm = TRUE) == 0) {
            tmp$inhibition[1] <- tmp$inhibition[1] - 10^-10
        }
        tmp.model <- drm(inhibition ~ dose, data = tmp, fct = L.4(fixed = c(NA, 
            tmp.min, Emax, NA)), na.action = na.omit)
        tmp$fitted.inhibition <- suppressWarnings(fitted(tmp.model))
        if (tmp$fitted.inhibition[nrow(response.mat) - 1] < 0) 
            tmp$fitted.inhibition[nrow(response.mat) - 1] <- tmp.min
        updated.col.mat[c(2:nrow(response.mat)), i] <- tmp$fitted.inhibition
    }
    updated.row.mat <- updated.single.mat
    for (i in 2:nrow(response.mat)) {
        tmp <- as.data.frame(mat.or.vec(ncol(response.mat) - 
            1, 0))
        tmp$dose <- as.numeric(colnames(response.mat)[-1])
        tmp$inhibition <- response.mat[i, c(2:ncol(response.mat))]
        tmp.min <- updated.single.mat[i, 1]
        if (var(tmp$inhibition, na.rm = TRUE) == 0) {
            tmp$inhibition[1] <- tmp$inhibition[1] - 10^-10
        }
        tmp.model <- drm(inhibition ~ dose, data = tmp, fct = L.4(fixed = c(NA, 
            tmp.min, Emax, NA)), na.action = na.omit)
        tmp$fitted.inhibition <- suppressWarnings(fitted(tmp.model))
        if (tmp$fitted.inhibition[ncol(response.mat) - 1] < 0) 
            tmp$fitted.inhibition[ncol(response.mat) - 1] <- tmp.min
        updated.row.mat[i, c(2:ncol(response.mat))] <- tmp$fitted.inhibition
    }
    fitted.mat <- (updated.col.mat + updated.row.mat)/2
    zip.mat <- updated.single.mat
    for (i in 2:nrow(updated.single.mat)) {
        for (j in 2:ncol((updated.single.mat))) {
            zip.mat[i, j] <- updated.single.mat[i, 1] + updated.single.mat[1, 
                j] - updated.single.mat[i, 1] * updated.single.mat[1, 
                j]/100
        }
    }
    fitted.mat[1, 1] <- 0
    zip.mat[1, 1] <- 0
    fitted.mat <- apply(fitted.mat, c(1, 2), function(x) ifelse(x > 
        100, 100, x))
    delta.mat <- (fitted.mat - zip.mat)
    delta.mat
}

```
here is mod of ZIP with trycatch. Fails by priting out NA for every single drug comb. 
```{r}
ZIP <- function (response.mat,correction, Emin = NA, Emax = NA,nan.handle = c("LL4", "L4")) {
# scores <- list()
#  method <- "Loewe"
  out <- tryCatch(ZIP_modded(response.mat,correction = correction, Emin = Emin, Emax = Emax, 
    nan.handle = nan.handle), error = function(e) NA)
#  {
#  x$method <- method
#  x$scores <- x$dose.response.mats
#  x$scores[!is.na(x$scores)] <- NA
#  x$scores <- replace(x$scores, !is.na(x$scores), NA)
#  purrr::map(!is.na(x$scores), NA)
#  x$scores <- lapply(x$scores, function(z) ifelse(!is.na(z), NA, z))
#  return(x)
#  })
  return(out)
}
```
BLISS
```{r}
Bliss <- function (response.mat = response.mat,correction = correction, Emin = NA, Emax = NA,nan.handle = c("LL4", "L4")) {
# scores <- list()
#  method <- "ZIP"
  out <- tryCatch(BLISS_modded(response.mat,correction = correction, Emin = Emin, Emax = Emax, 
    nan.handle = nan.handle), error = function(e) NA)
#  {
#  x$method <- method
#  x$scores <- x$dose.response.mats
#  x$scores[!is.na(x$scores)] <- NA
#  x$scores <- replace(x$scores, !is.na(x$scores), NA)
#  purrr::map(!is.na(x$scores), NA)
#  x$scores <- lapply(x$scores, function(z) ifelse(!is.na(z), NA, z))
#  return(x)
#  })
  return(out)
}
```
BLISS_modded 
```{r}
BLISS_modded<- function (response.mat, correction = TRUE, Emin = NA, Emax = NA, 
    nan.handle = c("LL4", "L4")) 
{
    if (correction) {
        response.mat <- BaselineCorrectionSD(response.mat, Emin = Emin, 
            Emax = Emax, nan.handle)$corrected.mat
    }
    drug1.response <- response.mat[, 1]
    drug2.response <- response.mat[1, ]
    ref.mat <- response.mat
    for (i in 2:nrow(response.mat)) {
        for (j in 2:ncol(response.mat)) {
            ref.mat[i, j] <- drug1.response[i] + drug2.response[j] - 
                drug1.response[i] * drug2.response[j]/100
        }
    }
    syn.mat <- response.mat - ref.mat
    syn.mat
}

```
HSA
```{r}
HSA <- function (response.mat = response.mat,correction = correction, Emin = NA, Emax = NA,nan.handle = c("LL4", "L4")) {
# scores <- list()
#  method <- "ZIP"
  out <- tryCatch(HSA_modded(response.mat,correction = correction, Emin = Emin, Emax = Emax, 
    nan.handle = nan.handle), error = function(e) NA)
#  {
#  x$method <- method
#  x$scores <- x$dose.response.mats
#  x$scores[!is.na(x$scores)] <- NA
#  x$scores <- replace(x$scores, !is.na(x$scores), NA)
#  purrr::map(!is.na(x$scores), NA)
#  x$scores <- lapply(x$scores, function(z) ifelse(!is.na(z), NA, z))
#  return(x)
#  })
  return(out)
}
```
HSA_modded
```{r}
HSA_modded <-function (response.mat, correction = TRUE, Emin = NA, Emax = NA, 
    nan.handle = c("LL4", "L4")) 
{
    if (correction) {
        response.mat <- BaselineCorrectionSD(response.mat, Emin = Emin, 
            Emax = Emax, nan.handle)$corrected.mat
    }
    drug1.response <- response.mat[, 1]
    drug2.response <- response.mat[1, ]
    ref.mat <- response.mat
    for (i in 2:nrow(response.mat)) {
        for (j in 2:ncol(response.mat)) {
            ref.mat[i, j] <- ifelse(drug1.response[i] > drug2.response[j], 
                drug1.response[i], drug2.response[j])
        }
    }
    syn.mat <- response.mat - ref.mat
    syn.mat
}

```

files were saved using 
`sink()`, `print(with max option set)`, `sink()`
and using simple `save()` to rds

original calculate synergy and zip functions
```{r}
CalculateSynergy <- function (data, method = "ZIP", correction = TRUE, Emin = 0, 
    Emax = 100, nan.handle = c("LL4", "L4")) 
{
    if (!is.list(data)) {
        stop("Input data is not a list format!")
    }
    if (!method %in% c("ZIP", "HSA", "Bliss", "Loewe")) {
        stop("The method parameter can only be one of the following: ZIP, HSA, Bliss and Loewe.")
    }
    dose.response.mats <- data$dose.response.mats
    num.pairs <- length(dose.response.mats)
    scores <- list()
    nan.handle <- match.arg(nan.handle)
    for (i in 1:num.pairs) {
        response.mat <- dose.response.mats[[i]]
        scores[[i]] <- switch(method, ZIP = ZIP(response.mat, 
            correction, Emin = Emin, Emax = Emax, nan.handle), 
            HSA = HSA(response.mat, correction, Emin = Emin, 
                Emax = Emax, nan.handle), Bliss = Bliss(response.mat, 
                correction, Emin = Emin, Emax = Emax, nan.handle), 
            Loewe = Loewe(response.mat, correction, Emin = Emin, 
                Emax = Emax, nan.handle))
    }
    data$scores <- scores
    data$method <- method
    return(data)
}
ZIP <- function (response.mat, correction = TRUE, Emin = 0, Emax = 100, 
    nan.handle = c("LL4", "L4")) 
{
    if (correction) {
        nan.handle <- match.arg(nan.handle)
        response.mat <- BaselineCorrectionSD(response.mat, NA, 
            NA, nan.handle)$corrected.mat
    }
    single.fitted <- FittingSingleDrug(response.mat, fixed = c(NA, 
        Emin, Emax, NA), nan.handle)
    drug.col.response <- single.fitted$drug.col.fitted
    drug.row.response <- single.fitted$drug.row.fitted
    updated.single.mat <- mat.or.vec(nrow(response.mat), ncol(response.mat))
    colnames(updated.single.mat) <- colnames(response.mat)
    rownames(updated.single.mat) <- rownames(response.mat)
    updated.single.mat[1, c(2:ncol(response.mat))] <- drug.col.response
    updated.single.mat[c(2:nrow(response.mat)), 1] <- drug.row.response
    updated.col.mat <- updated.single.mat
    for (i in 2:ncol(response.mat)) {
        tmp <- as.data.frame(mat.or.vec(nrow(response.mat) - 
            1, 0))
        tmp$dose <- as.numeric(rownames(response.mat)[-1])
        tmp$inhibition <- response.mat[c(2:nrow(response.mat)), 
            i]
        tmp.min <- updated.single.mat[1, i]
        if (var(tmp$inhibition, na.rm = TRUE) == 0) {
            tmp$inhibition[1] <- tmp$inhibition[1] - 10^-10
        }
        tmp.model <- drm(inhibition ~ dose, data = tmp, fct = L.4(fixed = c(NA, 
            tmp.min, Emax, NA)), na.action = na.omit)
        tmp$fitted.inhibition <- suppressWarnings(fitted(tmp.model))
        if (tmp$fitted.inhibition[nrow(response.mat) - 1] < 0) 
            tmp$fitted.inhibition[nrow(response.mat) - 1] <- tmp.min
        updated.col.mat[c(2:nrow(response.mat)), i] <- tmp$fitted.inhibition
    }
    updated.row.mat <- updated.single.mat
    for (i in 2:nrow(response.mat)) {
        tmp <- as.data.frame(mat.or.vec(ncol(response.mat) - 
            1, 0))
        tmp$dose <- as.numeric(colnames(response.mat)[-1])
        tmp$inhibition <- response.mat[i, c(2:ncol(response.mat))]
        tmp.min <- updated.single.mat[i, 1]
        if (var(tmp$inhibition, na.rm = TRUE) == 0) {
            tmp$inhibition[1] <- tmp$inhibition[1] - 10^-10
        }
        tmp.model <- drm(inhibition ~ dose, data = tmp, fct = L.4(fixed = c(NA, 
            tmp.min, Emax, NA)), na.action = na.omit)
        tmp$fitted.inhibition <- suppressWarnings(fitted(tmp.model))
        if (tmp$fitted.inhibition[ncol(response.mat) - 1] < 0) 
            tmp$fitted.inhibition[ncol(response.mat) - 1] <- tmp.min
        updated.row.mat[i, c(2:ncol(response.mat))] <- tmp$fitted.inhibition
    }
    fitted.mat <- (updated.col.mat + updated.row.mat)/2
    zip.mat <- updated.single.mat
    for (i in 2:nrow(updated.single.mat)) {
        for (j in 2:ncol((updated.single.mat))) {
            zip.mat[i, j] <- updated.single.mat[i, 1] + updated.single.mat[1, 
                j] - updated.single.mat[i, 1] * updated.single.mat[1, 
                j]/100
        }
    }
    fitted.mat[1, 1] <- 0
    zip.mat[1, 1] <- 0
    fitted.mat <- apply(fitted.mat, c(1, 2), function(x) ifelse(x > 
        100, 100, x))
    delta.mat <- (fitted.mat - zip.mat)
    delta.mat
}

```
##### what actually worked is to have 
`today.full.ZIP <- CalculateSynergy(reshaped)` 
todawithout any parameters

## aug 8.08
get first two of each of the nested lists. Test
```{r}
first.two.response <- today.ZIP$dose.response.mats[c(1,2)]
first.two.drugs <- today.ZIP$drug.pairs[c(1,2),]
first.two.scores <- today.ZIP$scores[c(1,2)]
first.two.method <- today.ZIP$method
first.two <- list()
first.two$dose.response.mats <- first.two.response
first.two$scores <- first.two.scores
first.two$drugs <- first.two.drugs
first.two$method <- first.two.method
first.two.response.Bliss <- today.Bliss$dose.response.mats[c(1,2)]
first.two.drugs.Bliss <- today.Bliss$drug.pairs[c(1,2),]
first.two.scores.Bliss <- today.Bliss$scores[c(1,2)]
first.two.method.Bliss <- today.Bliss$method
first.two.Bliss <- list()
first.two.Bliss$dose.response.mats <- first.two.response.Bliss
first.two.Bliss$scores <- first.two.scores.Bliss
first.two.Bliss$drugs <- first.two.drugs.Bliss
first.two.Bliss$method <- first.two.method.Bliss
```
libraries for melting and joining
```{r}
library(reshape2) #cast/melt
library(tidyverse) #spread/gather
library(rlist) #list join
options(error = recover)
```
checking the melting for ZIP.Test
```{r}
x.ZIP <- melt(first.two$dose.response.mats)
y.ZIP <- melt(first.two$scores)
xy.ZIP <- merge(x.ZIP,y.ZIP, by = c('Var1', 'Var2', 'L1'))
```
checking the melting for Bliss.Test
```{r}
x.Bliss <- melt(first.two.Bliss$dose.response.mats)
y.Bliss <- melt(first.two.Bliss$scores)
xy.Bliss <- merge(x.Bliss,y.Bliss, by = c('Var1', 'Var2', 'L1'))
```
general rename for ZIP.Test
```{r}
colnames(xy.ZIP)[which(names(xy.ZIP) == "Var1")] <- "ConcR"
colnames(xy.ZIP)[which(names(xy.ZIP) == "Var2")] <- "ConcC"
colnames(xy.ZIP)[which(names(xy.ZIP) == "L1")] <- "blockIDs"
colnames(xy.ZIP)[which(names(xy.ZIP) == "value.x")] <- "Response_inhibition"
synergy.type.ZIP <- paste('Synergy', first.two$method, sep = '_')
colnames(xy.ZIP)[which(names(xy.ZIP) == "value.y")] <- synergy.type.ZIP
```
general rename for Bliss.Test
```{r}
colnames(xy.Bliss)[which(names(xy.Bliss) == "Var1")] <- "ConcR"
colnames(xy.Bliss)[which(names(xy.Bliss) == "Var2")] <- "ConcC"
colnames(xy.Bliss)[which(names(xy.Bliss) == "L1")] <- "blockIDs"
colnames(xy.Bliss)[which(names(xy.Bliss) == "value.x")] <- "Response_inhibition"
synergy.type.Bliss <- paste('Synergy', first.two.Bliss$method, sep = '_')
colnames(xy.Bliss)[which(names(xy.Bliss) == "value.y")] <- synergy.type.Bliss
```

merging drug names.Test
```{r}
first.two.no.celllines <- merge(xy.ZIP, first.two$drugs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
first.two.no.celllines.Bliss <- merge(xy.Bliss, first.two.Bliss$drugs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)

first.two.no.celllines.ZIP.Bliss <- merge(first.two.no.celllines, first.two.no.celllines.Bliss, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
```
prep for monstrous loop
```{r}
first.two.response <- today.ZIP$dose.response.mats
first.two.drugs <- today.ZIP$drug.pairs
first.two.scores <- today.ZIP$scores
first.two.method <- today.ZIP$method
first.two <- list()
first.two$dose.response.mats <- first.two.response
first.two$scores <- first.two.scores
first.two$drugs <- first.two.drugs
first.two$method <- first.two.method

first.two.response.Bliss <- today.Bliss$dose.response.mats
first.two.drugs.Bliss <- today.Bliss$drug.pairs
first.two.scores.Bliss <- today.Bliss$scores
first.two.method.Bliss <- today.Bliss$method
first.two.Bliss <- list()
first.two.Bliss$dose.response.mats <- first.two.response.Bliss
first.two.Bliss$scores <- first.two.scores.Bliss
first.two.Bliss$drugs <- first.two.drugs.Bliss
first.two.Bliss$method <- first.two.method.Bliss

first.two.response.L <- today.Loewe$dose.response.mats
first.two.drugs.L <- today.Loewe$drug.pairs
first.two.scores.L <- today.Loewe$scores
first.two.method.L <- today.Loewe$method
first.two.L <- list()
first.two.L$dose.response.mats <- first.two.response.L
first.two.L$scores <- first.two.scores.L
first.two.L$drugs <- first.two.drugs.L
first.two.L$method <- first.two.method.L

first.two.response.H <- today.HSA$dose.response.mats
first.two.drugs.H <- today.HSA$drug.pairs
first.two.scores.H <- today.HSA$scores
first.two.method.H <- today.HSA$method
first.two.H <- list()
first.two.H$dose.response.mats <- first.two.response.H
first.two.H$scores <- first.two.scores.H
first.two.H$drugs <- first.two.drugs.H
first.two.H$method <- first.two.method.H
```
treating NAs (when taking blockID 226).Test
```{r}
#colnames(first.two$dose.response.mats[[2]])
#rownames(first.two$dose.response.mats[[2]])
#dim(first.two$dose.response.mats[[2]])
cols <- colnames(first.two$dose.response.mats[[which(is.na(first.two$scores))]])
rows <- rownames(first.two$dose.response.mats[[which(is.na(first.two$scores))]])
numrows <- dim(first.two$dose.response.mats[[which(is.na(first.two$scores))]])[1]
numcols <- dim(first.two$dose.response.mats[[which(is.na(first.two$scores))]])[2]

first.two$scores[[which(is.na(first.two$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))


```
fuck it, let's do it with loops yo
```{r}
datalist = list()

for (i in 1:nrow(first.two$drugs))
{
  a <- list()
  b <- list()
  c <- list()
  d <- list()
  
  
  a$dose.response.mats <- first.two$dose.response.mats[i]
  a$scores <- first.two$scores[i]
  a$method <- first.two$method
  a$drugs <- first.two$drugs[i,]
  
    if (is.na(a$scores)) {
    cols <- colnames(a$dose.response.mats[[which(is.na(a$scores))]])
    rows <- rownames(a$dose.response.mats[[which(is.na(a$scores))]])
    numrows <- dim(a$dose.response.mats[[which(is.na(a$scores))]])[1]
    numcols <- dim(a$dose.response.mats[[which(is.na(a$scores))]])[2]
    a$scores[[which(is.na(a$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }
  
  b$dose.response.mats <- first.two.Bliss$dose.response.mats[i]
  b$scores <- first.two.Bliss$scores[i]
  b$method <- first.two.Bliss$method
  b$drugs <- first.two.Bliss$drugs[i,]
  
    if (is.na(b$scores)) {
    cols <- colnames(b$dose.response.mats[[which(is.na(b$scores))]])
    rows <- rownames(b$dose.response.mats[[which(is.na(b$scores))]])
    numrows <- dim(b$dose.response.mats[[which(is.na(b$scores))]])[1]
    numcols <- dim(c$dose.response.mats[[which(is.na(b$scores))]])[2]
    b$scores[[which(is.na(b$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }
  
  c$dose.response.mats <- first.two.L$dose.response.mats[i]
  c$scores <- first.two.L$scores[i]
  c$method <- first.two.L$method
  c$drugs <- first.two.L$drugs[i,]
  
    if (is.na(c$scores)) {
      
    cols <- colnames(c$dose.response.mats[[which(is.na(c$scores))]])
    rows <- rownames(c$dose.response.mats[[which(is.na(c$scores))]])
    numrows <- dim(c$dose.response.mats[[which(is.na(c$scores))]])[1]
    numcols <- dim(c$dose.response.mats[[which(is.na(c$scores))]])[2]
    c$scores[[which(is.na(c$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }
  
  d$dose.response.mats <- first.two.H$dose.response.mats[i]
  d$scores <- first.two.H$scores[i]
  d$method <- first.two.H$method
  d$drugs <- first.two.H$drugs[i,]
  
    if (is.na(d$scores)) {
    cols <- colnames(d$dose.response.mats[[which(is.na(d$scores))]])
    rows <- rownames(d$dose.response.mats[[which(is.na(d$scores))]])
    numrows <- dim(d$dose.response.mats[[which(is.na(d$scores))]])[1]
    numcols <- dim(d$dose.response.mats[[which(is.na(d$scores))]])[2]
    d$scores[[which(is.na(d$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }

  x.ZIP <- melt(a$dose.response.mats)
  y.ZIP <- melt(a$scores)
  xy.ZIP <- merge(x.ZIP,y.ZIP, by = c('Var1', 'Var2', 'L1'))
  
  x.Bliss <- melt(b$dose.response.mats)
  y.Bliss <- melt(b$scores)
  xy.Bliss <- merge(x.Bliss,y.Bliss, by = c('Var1', 'Var2', 'L1'))
  
  x.Loewe <- melt(c$dose.response.mats)
  y.Loewe <- melt(c$scores)
  xy.Loewe <- merge(x.Loewe,y.Loewe, by = c('Var1', 'Var2', 'L1'))
  
  x.HSA <- melt(d$dose.response.mats)
  y.HSA <- melt(d$scores)
  xy.HSA <- merge(x.HSA,y.HSA, by = c('Var1', 'Var2', 'L1'))
  
  colnames(xy.ZIP)[which(names(xy.ZIP) == "Var1")] <- "ConcR"
  colnames(xy.ZIP)[which(names(xy.ZIP) == "Var2")] <- "ConcC"
  colnames(xy.ZIP)[which(names(xy.ZIP) == "L1")] <- "blockIDs"
  xy.ZIP$blockIDs <- i
  colnames(xy.ZIP)[which(names(xy.ZIP) == "value.x")] <- "Response_inhibition"
  synergy.type.ZIP <- paste('Synergy', first.two$method, sep = '_')
  colnames(xy.ZIP)[which(names(xy.ZIP) == "value.y")] <- synergy.type.ZIP
  
  colnames(xy.Bliss)[which(names(xy.Bliss) == "Var1")] <- "ConcR"
  colnames(xy.Bliss)[which(names(xy.Bliss) == "Var2")] <- "ConcC"
  colnames(xy.Bliss)[which(names(xy.Bliss) == "L1")] <- "blockIDs"
  xy.Bliss$blockIDs <- i
  colnames(xy.Bliss)[which(names(xy.Bliss) == "value.x")] <- "Response_inhibition"
  synergy.type.Bliss <- paste('Synergy', first.two.Bliss$method, sep = '_')
  colnames(xy.Bliss)[which(names(xy.Bliss) == "value.y")] <- synergy.type.Bliss
  
  colnames(xy.Loewe)[which(names(xy.Loewe) == "Var1")] <- "ConcR"
  colnames(xy.Loewe)[which(names(xy.Loewe) == "Var2")] <- "ConcC"
  colnames(xy.Loewe)[which(names(xy.Loewe) == "L1")] <- "blockIDs"
  xy.Loewe$blockIDs <- i
  colnames(xy.Loewe)[which(names(xy.Loewe) == "value.x")] <- "Response_inhibition"
  synergy.type.Loewe <- paste('Synergy', first.two.L$method, sep = '_')
  colnames(xy.Loewe)[which(names(xy.Loewe) == "value.y")] <- synergy.type.Loewe
  
  colnames(xy.HSA)[which(names(xy.HSA) == "Var1")] <- "ConcR"
  colnames(xy.HSA)[which(names(xy.HSA) == "Var2")] <- "ConcC"
  colnames(xy.HSA)[which(names(xy.HSA) == "L1")] <- "blockIDs"
  xy.HSA$blockIDs <- i
  colnames(xy.HSA)[which(names(xy.HSA) == "value.x")] <- "Response_inhibition"
  synergy.type.HSA <- paste('Synergy', first.two.H$method, sep = '_')
  colnames(xy.HSA)[which(names(xy.HSA) == "value.y")] <- synergy.type.HSA
  
  
  first.two.no.celllines <- merge(xy.ZIP, a$drugs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  first.two.no.celllines.Bliss <- merge(xy.Bliss, b$drugs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  first.two.no.celllines.Loewe <- merge(xy.Loewe, c$drugs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  first.two.no.celllines.HSA <- merge(xy.HSA, d$drugs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  
  first.two.no.celllines.ZIP.Bliss <- merge(first.two.no.celllines, first.two.no.celllines.Bliss, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
  first.two.no.celllines.ZIP.Bliss.Loewe <- merge(first.two.no.celllines.ZIP.Bliss, first.two.no.celllines.Loewe, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
  first.two.no.celllines.ZIP.Bliss.Loewe.HSA <- merge(first.two.no.celllines.ZIP.Bliss.Loewe, first.two.no.celllines.HSA, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
  
  
  
  datalist[[i]] <- first.two.no.celllines.ZIP.Bliss.Loewe.HSA
  
}
big_data = do.call(rbind, datalist)
```

adding cell lines
```{r}
asdf <- input.sample %>%
  group_by(BlockID) %>%
    transmute(cell_line) %>%
      summarise(cell_line = unique(cell_line))

testing <- big_data
out <- merge(testing, asdf, by.x=c('blockIDs'),by.y=c('BlockID'),all.x = T )
write.csv(out,file = '~/Documents/fimm_files/synergy_calc_august/output_alls.csv')
```














