---
title: "nci_Aug"
output: html_notebook
---
##### setting up 
```{r}
library("tidyverse")
library("synergyfinder")
setwd('/Users/zagidull/Documents/fimm_files/synergy_calc_august/almanac')
```

##### reading in the data
```{r}
nov2017 <- read_csv('ComboDrugGrowth_Nov2017.csv', progress = F) # it is nice to read as tibble, this helps get rid of formatting errors
nov2017$TESTDATE <- as.Date(nov2017$TESTDATE,"%m/%d/%Y") #correcting DATA col class
sapply(nov2017, class) #to check for correctedness of setting the classes
```

##### sort by screener and remove NCI

`A total of 5,232 drug pairs were evaluated in each of the cell lines; 304,549 experiments were performed to test each drug at either 9 or 15 combination dose points, for a total of 2,809,671 dose combinations`

##### here we take one cell line and prepare it for processing. 
```{r}
for_sorting <- nov2017

for_sorting_one <- for_sorting %>% # taking one cell line as an example
  filter(CELLNAME=='786-0')

saveRDS(for_sorting_one,"for_sorting_one" )
```

##### starting work. As a result there are two files: singles and combos, where singles have single drugs testd. Combos are obv drug pairs
```{r}
for_sorting_one$groupID <- for_sorting_one %>% 
  group_by(NSC1,NSC2) %>%
    group_indices(.)

for_sorting_one <- for_sorting_one %>%
  group_by(groupID)

before <- colnames(for_sorting_one)
for_sorting_one <- for_sorting_one[vapply(for_sorting_one, function(x) length(unique(x)) > 1, logical(1L))] # drop all columns with a single unique value


for_sorting_one <- for_sorting_one[,-c(22,21,20,19,18,17)] # drop useless 
for_sorting_one <- for_sorting_one[, -c(17)]
after <- colnames(for_sorting_one)

dropped <- setdiff(before, after)

singles <- for_sorting_one %>% # this is for al lthe cases when only a single drug was tested
  filter(CONCINDEX2 == 0)

combos <- for_sorting_one %>% # this is for al lthe cases when two drugs were tested
  filter(CONCINDEX2 != 0)
```

##### we need to make sure that we have the same drugID (NSC numbers) in both singles and combos
```{r}
setdiff(singles$NSC1, combos$NSC1)
NSC_to_drop <- setdiff(singles$NSC1, combos$NSC1)
# lets get rid of the NSC that was tested only as single
singles <- singles[-which(singles$NSC1 == NSC_to_drop),]
print(setdiff(singles$NSC1, combos$NSC1))
for_syn_calc <- rbind(singles, combos)
saveRDS(for_syn_calc, "for_syn_calc")

```
##### now let's group by NSC1 to do the reshape and synergy calc
```{r}
nrow(for_syn_calc)
for_syn_calc <- for_syn_calc[-which(is.na(for_syn_calc$PREFIX2)), ] # because all values from the ALMANAC study have prefix S
nrow(for_syn_calc)

for_syn_calc$groups <- for_syn_calc %>% # i m not sure what s the point of doing that anymore. I dont really need that in R. It is here in case of applying python add_col_row.py
  group_by(NSC1) %>%
    group_indices(.)
for_syn_calc <- for_syn_calc %>%
  group_by(groups) %>%
    mutate(num_rows = n()) %>%
      arrange(groups)
      

for_syn_calc <- for_syn_calc[vapply(for_syn_calc, function(x) length(unique(x)) > 1, logical(1L))] # removing vars that have one single unique value throughout the whole dataset

withMeanGrowth <- for_syn_calc %>%
  group_by(groups) %>%
      group_by(CONCINDEX1,CONC1, CONC2, NSC1, NSC2) %>%
        mutate(mean_single = mean(PERCENTGROWTH)) %>% # this gets mean values for all the experiments where one drug is tested by itself in cell line == 786-0
          arrange(groups)

withMeanGrowth_noDups <- withMeanGrowth[!duplicated(withMeanGrowth[c("NSC1","CONC1", "CONC2", "mean_single")]),] # i don't need to have duplicated rows (if these are indeed duplicated experiments) after doing mutate(mean_single=...)

#so test and test2 are identical in terms of unqiue info, but test2 has all the duplicated rows dropped. Here are there (FYI)
dropped_for786_0 <- setdiff(withMeanGrowth,withMeanGrowth_noDups)

withMeanGrowth_noDups_cleanedCols <- withMeanGrowth_noDups[,-c(15,16,2,3,4,5)] # dropping useless columns

uniqueDrug1Drub2Combos <- withMeanGrowth_noDups_cleanedCols %>%
  ungroup() %>%
  mutate(NSC1,NSC2) %>%
  select(NSC1, NSC2) %>%
     unique # this gives a list of unique combinations whereby each combination should result in a separate group for synergyfinder::reshape()
```
##### testing transformation
```{r}
oneComboTest_740_737754 <- withMeanGrowth_noDups %>%
  filter((NSC1 == 740 & NSC2 ==737754) | (NSC1 == 740 & is.na(NSC2)) | (NSC1 == 737754 & is.na(NSC2)) | (NSC1 == 737754 & NSC2 ==740)) #%>%
#    mutate(identifier = "id1")

oneComboTest_740_754143 <- withMeanGrowth_noDups %>%
  filter((NSC1 == 740 & NSC2 ==754143) | (NSC1 == 740 & is.na(NSC2)) | (NSC1 == 754143 & is.na(NSC2)) | (NSC1 == 754143 & NSC2 ==740)) #%>%
#    mutate(identifier = "id2")

rowInsert <- function(x, blockID){ #adding zero row
  class_old <- sapply(x, class)
#  print(class_old)
  NSC2_un <- unique(x$NSC2[!is.na(x$NSC2)])
  
  predicate <- function(x) {
    if (x[["CONCINDEX2"]]== 0 & x[["NSC1"]] == NSC2_un) {
      return(T)
    }
    else {return(F)}
  }
  
  exchange <- function(x) {
    if (predicate(x)){
    placeholderIndex <- x[["CONCINDEX1"]]
    placeholderConc <- x[["CONC1"]]
    x[["CONCINDEX1"]] <- x[["CONCINDEX2"]]
    x[["CONCINDEX2"]] <- placeholderIndex
    x[["CONC1"]] <- x[["CONC2"]]
    x[["CONC2"]] <- placeholderConc
    return(x)
    } else { return(x)}
  }  
  x <- as.tibble(t(apply(x, 1, exchange)))
  x[] <- Map(`class<-`,x, class_old) #brilliant! https://stackoverflow.com/questions/27435873/changing-class-of-data-frame-columns-using-strings
#  class_new <- sapply(x, class)
#  print(class_new)
  
  
  
  
  NSC1 <- dplyr::setdiff(unique(x$NSC1[!is.na(x$NSC1)]),NSC2_un)
#  NSC1 <- unique(x$NSC1[!is.na(x$NSC1)])
  x[(nrow(x)+1),] <- x[nrow(x),]
  x[nrow(x),]$NSC1 <- NSC1
  x[nrow(x),]$NSC2 <- NSC2_un
  x[nrow(x),]$CONCINDEX1 <- x[nrow(x),]$CONCINDEX2 <- x[nrow(x),]$CONC1 <- x[nrow(x),]$CONC2 <- 0
  x[nrow(x),]$mean_single <- 0
  x[nrow(x),]$SAMPLE2 <- unique(x$SAMPLE2[!is.na(x$SAMPLE2)])

  x$Row <- x$CONCINDEX1 + 1
  x$Col <- x$CONCINDEX2 + 1
  x$BlockID <- c(blockID) # sub with cell-line
  x$Response <- x$mean_single
  x$ConcRowUnit <- x$ConcColUnit <- c("M")
  x$ConcRow <- x$CONC1
  x$ConcCol <- x$CONC2
  x$ConcCol[is.na(x$ConcCol) ] <- 0
  x$DrugRow <- NSC1
  x$DrugCol <- NSC2_un
  x[is.na(x$ConcRow),]$ConcRow <- 0
  x <- x[,-c(1:17)]
  return(x)
}
```
##### testing on two
```{r}
twoblocks <- rbind(oneComboTest_740_3053, oneComboTest_740_754143)
meow <- twoblocks %>%
  group_by(identifier) %>%
    do(rowInsert(., 'test'))
```

##### testing on all the celllines
```{r}
uniqueDrug1Drub2CombosNoNA <-  uniqueDrug1Drub2Combos[!is.na(uniqueDrug1Drub2Combos$NSC2),] # no na's
uniqueDrug1Drub2CombosNoNA[c(100:nrow(uniqueDrug1Drub2CombosNoNA)),] -> test
#datalist = list() dont do it !!!

#assembly <- function(x,y) {
#  sequence <- nrow(y)
#  
#  
#  output <- list()
  datalist_full <- list()
  x= withMeanGrowth_noDups
  y = uniqueDrug1Drub2CombosNoNA
  for (i in 1:nrow(y)) {
    perrow <- list()
    input <- list()
    output1 <- list()
    param1 <- vector(mode="numeric", length=1)
    param2 <- vector(mode="numeric", length=1)
    perrow<- y[i,]

    param1 <- as.numeric(perrow[1])
    param2 <- as.numeric(perrow[2])

    input <- filter(x, ((NSC1 == param1 & NSC2 == param2) | (NSC1 == param1 & is.na(NSC2)) | (NSC1 == param2 & is.na(NSC2)) | (NSC1 == param2 & NSC2 ==param1)))
    input[,ncol(input)+1] <- i
    
    output1 <-  tryCatch(rowInsert(input,i), error = function(e) NA)
#    output1 <- rowInsert(input,i)
    datalist_full[[i]] <- output1

  }
big_data_full = do.call(rbind, datalist_full)

#}
#assembly(withMeanGrowth_noDups, test)

#apply(withMeanGrowth_noDups,1 )

#by(test, 1:nrow(test), function(row) print(test[row,]))
```


##### this is a graveyard of ideas. Lasciate ogni speranza voi ch'entrante
```{r}

test7 <- test6[test6$NSC1==752,]
test7 %>%
  rowwise() %>%
    filter((test5$NSC1 == NSC1 & is.na(test5$NSC2)) | (test5$NSC1 == NSC2 & is.na(test5$NSC1)) | (test5$NSC1 == NSC2 & test5$NSC2 == NSC1) | (test5$NSC1 == NSC1 & test5$NSC2 == NSC2))
    



oneComboTest <- test2 %>%
  filter((NSC1 == 740 & NSC2 ==53) | (NSC1 == 740 & is.na(NSC2)) | (NSC1 == 53 & is.na(NSC2)) | (NSC1 == 53 & NSC2 ==740))

for_show <- test %>%
  filter((NSC1 == 740 & NSC2 ==3053) | (NSC1 == 740 & is.na(NSC2)) | (NSC1 == 3053 & is.na(NSC2)) | (NSC1 == 3053 & NSC2 ==740))
for_show1 <- for_show[vapply(for_show, function(x) length(unique(x)) > 1, logical(1L))]

test4 <- test3[,-c(1,2,3,4,5,7,11,16)]






test1 <- for_syn_calc %>%
  group_by(groups) %>%
    filter(CONCINDEX2 == 0) %>%
      group_by(CONCINDEX1, NSC1) %>%
        mutate(mean_single = mean(PERCENTGROWTH))




out <- for_syn_calc %>%
  filter((NSC1 == 752 & NSC2 ==3088) | (NSC1 == 752 & is.na(NSC2)) | (NSC1 == 3088 & is.na(NSC2)) | (NSC1 == 3088 & NSC2 ==752))
out <- out[vapply(out, function(x) length(unique(x)) > 1, logical(1L))]
out <- out[, -c(22,19,18,17,16)] # in case not dropping in the previous block of code

out_mean <- out %>%
#  group_by(groupID) %>%
 #   filter(groupID != 134) %>%
      filter((CONCINDEX1 == 1 & CONCINDEX2 == 0 )| (CONCINDEX1 == 2 & CONCINDEX2 == 0 ) | (CONCINDEX1 == 3 & CONCINDEX2 == 0 ) |  (CONCINDEX2 != 0 )) %>%
        group_by(NSC1, NSC2,CONCINDEX1,CONC1, CONC2) %>%
#          group_by(groupID) %>%
            transmute(mean_single = mean(PERCENTGROWTH)) 

out_mean %>% 
  group_by(groupID, NSC1, CONC1, CONC2) %>%
    transmute(mean_single)

for_syn_calc %>%
  group_by(groups) %>%
    summarise(unique(NSC1), unique(NSC2) )
```


##### Per me si va ne la citta dolente


```{r}
for_sorting_one$combo <- for_sorting_one %>%
  group_by(CONCINDEX2) %>%
    group_indices(.)


singles$combo <- singles %>% #here combo becomes groupID
  group_by(NSC1) %>%
    group_indices(.)

singles <- singles %>%
   group_by(combo)%>%
      mutate(num_rows = n())


for_sorting$groupID <- for_sorting %>% # each drug pair tested in each of the cell lines
  group_by(CELLNAME) %>%
    group_indices(.)




for_sorting <- for_sorting %>%
  group_by(CELLNAME) %>%
    mutate(num_rows = n())

six <- for_sorting %>%
  filter(num_rows == 6)
three <- for_sorting %>%
  filter(num_rows == 3)
nine <- for_sorting %>%
  filter(num_rows == 9)

test <- for_sorting %>% #this plate has both single drug in 5 doses and drug combo
    filter(NSC1 == 752 & NSC2 == 3088 & CELLNBR == 18)


for_sorting[!is.na(for_sorting$PERCENTGROWTH),]
```



##### per me si va ne l'etterno dolore,

##### prep for all where there is a complete 3x3 matrix with no NA. todo: try complete.case instad of na.omit
```{r}
noNA_all_rows <- nov2017 %>%
  na.omit() %>%
    group_by(NSC1, NSC2, SCREENER, PLATE, CELLNBR) %>%
      summarise(number = n()) # there are 300091 different groups

noNA_grouped <-nov2017 %>%
  na.omit() %>%
    group_by(NSC1, NSC2, PLATE, CELLNBR) %>% # there are 2700819 rows all having 3x3 matrices. Ok first set of cals
      
noNA_grouped$index <- noNA_grouped %>%  
  group_indices(.) # attach index column with index ID
```
##### per me si va tra la perduta gente

##### testing
```{r}
onegroup <- noNA_grouped %>%
  filter(index==1)
```
##### Giustizia mosse il mio alto fattore:

##### fucjuing around. Selecting on per-screener basis. Because in some 3x3 was tested, in others 5x3
```{r}
try_sort <- nov2017 # trying to figure out an effective screener way
try_sort$groups <- try_sort %>% #adding groups based on panelnbr, drug1, drug2, plateID and screener institute
  group_by(PANELNBR, NSC1, NSC2, PLATE, SCREENER, CELLNBR) %>%
    group_indices(.)
try_sort <- try_sort %>% # adding row count
  group_by(groups) %>%
    mutate(num_rows = n())
print(try_sort[(try_sort$SCREENER == "FG" & try_sort$CELLNBR==18 & try_sort$NSC1 == 752 & try_sort$CELLNAME == "786-0" & try_sort$PANELNBR == 9 & try_sort$PANEL == 'Renal Cancer'), ], n = 395)


withNA <- try_sort %>% # technically all drug combs have to have either 3x3 or 3x5 doses. So each not like that has NA as one of the values
  filter(num_rows !=9 & num_rows != 15)
noNA <- try_sort %>% # technically all drug combs have to have either 3x3 or 3x5 doses. NO na. Wait, so this has only num_rows == 9. Interesting. So 15 is not necessary that s a bit annoying coz there should be 
  filter(num_rows ==9 | num_rows == 3)

noNA_clean <- noNA[,-c(1,2,3,4,5,6,7,8,10,14,16,22,23,24,25,26,27,28,29)]
```
##### fecemi la divina potestate

##### let's do the reshaping
```{r}
library(synergyfinder)

```

##### la somma sapienza e 'l primo amore. 

##### this is more playing
```{r}

screener_select <- nov2017 # make copy
screener_select$by_screener <- nov2017 %>% #group_by screener and add indices
  group_by(SCREENER) %>%
    group_indices(.) 

all_rows_noNA <- all_rows %>%
  na.omit() 

all_rows_noNA$groups <- all_rows_noNA%>%
    group_by(NSC1, NSC2, SCREENER, PLATE, CELLNBR)%>%
      group_indices(.)
  
all_rows_noNA <- all_rows_noNA %>% # adding row count
  group_by(groups) %>%
    mutate(num_rows = n())

try_one_more <- nov2017
try_one_more$groups <- try_one_more %>% #adding groups based on panelnbr, drug1, drug2, plateID and screener institute
  group_by(NSC1, NSC2, PLATE, PANELNBR, CELLNBR ,SCREENER) %>%
    group_indices(.)
try_one_more <- try_one_more %>% # adding row count
  group_by(groups) %>%
    mutate(num_rows = n())
noNA_more <- try_one_more %>% # technically all drug combs have to have either 3x3 or 3x5 doses. NO na. Wait, so this has only num_rows == 9. Interesting
  filter(num_rows ==9)
unique(try_one_more$num_rows)

plate2998 <- try_sort %>% #this plate has both single drug in 5 doses and drug combo
  filter(PLATE == 2998) %>%
    filter(NSC1 == 256439 & NSC2 == 747972)


oneA <- screener_select %>% #select one screener lab
  filter(screener_select$by_screener==1) 

plate8137 <- oneA %>% #this plate has both single drug in 5 doses and drug combo
  filter(PLATE == 8137) %>%
    group_by(PANELNBR, NSC1, NSC2, PLATE, SCREENER)

cleanup_plate8137 <- plate8137[,-c(1,2,3,4,5,8,13,19,23,24,25)] #remove not used atm columns

  
FF <- screener_select %>%
  filter(screener_select$by_screener==2)
FG <- screener_select %>%
  filter(screener_select$by_screener==3)
```
##### Dinanzi a me non fuor cose create

##### something
```{r}
result <- nov2017 %>%
      filter(nov2017$NSC1==740 & nov2017$PLATE=="01_21" & nov2017$CELLNBR=='6' & is.na(nov2017$NSC2)) # is drugA, drugB and cellline enough to uniquely identify the combination? A: seems like it
result1 <- nov2017 %>%
      filter(nov2017$SAMPLE1==740 & ?pnov2017$NSC2==369100 & nov2017$CELLNAME=='NCI-H522') 

result1 <- nov2017 %>%
      filter(nov2017$COMBODRUGSEQ=='260496') 

nov2017[c(1:27),] -> nov2017short
result_tib <- as_tibble(nov2017short)
grouped_result <- nov2017 %>%
  group_by(PLATE, NSC1, NSC2) 

number_of_rows <- nov2017 %>%
  group_by(NSC1, NSC2,CELLNAME, PLATE) %>%
    summarise(number = n()) 

NSC2NA <- number_of_rows %>%
  filter(is.na(NSC2))
fourrow <- number_of_rows %>%
  filter(number==4)
```