---
title: "prep for database input"
output: html_notebook
---
## this is used to transform synergyfinder::CalculateSynergy() results into the format that could be loaded for the DB

## aug 28.08
get first two of each of the CalculateSynergy output files. For testing purposes. There are three input files.
NB: first.two.$ without the method means it is "ZIP". Since it is the default for the CalculateSynergy
```{r}
#create placeholder lists
first.two <- list()
first.two.Bliss <- list()
first.two.HSA <- list()

#populate for ZIP
first.two$dose.response.mats <- reshaped.ZIP$dose.response.mats[c(1,2)]
first.two$drug.pairs <- reshaped.ZIP$drug.pairs[c(1,2),]
first.two$scores <- reshaped.ZIP$scores[c(1,2)]
first.two$method <- reshaped.ZIP$method

#populate for Bliss
first.two.Bliss$dose.response.mats <- reshaped.Bliss$dose.response.mats[c(1,2)]
first.two.Bliss$drug.pairs <- reshaped.Bliss$drug.pairs[c(1,2),]
first.two.Bliss$scores <- reshaped.Bliss$scores[c(1,2)]
first.two.Bliss$method <- reshaped.Bliss$method

#populate for HSA
first.two.HSA$dose.response.mats <- reshaped.HSA$dose.response.mats[c(1,2)]
first.two.HSA$drug.pairs <- reshaped.HSA$drug.pairs[c(1,2),]
first.two.HSA$scores <- reshaped.HSA$scores[c(1,2)]
first.two.HSA$method <- reshaped.HSA$method
```
libraries for melting and joining
```{r}
library(reshape2) #cast/melt
library(tidyverse) #spread/gather
library(rlist) #list join
options(error = recover)
```
checking the melting for ZIP.Test
```{r}
x.ZIP <- melt(first.two$dose.response.mats)
y.ZIP <- melt(first.two$scores)
xy.ZIP <- merge(x.ZIP,y.ZIP, by = c('Var1', 'Var2', 'L1'))
```
checking the melting for Bliss.Test
```{r}
x.Bliss <- melt(first.two.Bliss$dose.response.mats)
y.Bliss <- melt(first.two.Bliss$scores)
xy.Bliss <- merge(x.Bliss,y.Bliss, by = c('Var1', 'Var2', 'L1'))
```
checking the melting for HSA.Test
```{r}
x.HSA <- melt(first.two.HSA$dose.response.mats)
y.HSA <- melt(first.two.HSA$scores)
xy.HSA <- merge(x.HSA,y.HSA, by = c('Var1', 'Var2', 'L1'))
```

creating a dictionary to substitute the values with
```{r}
#here comes the problem whereby i right now have a numerical + string blockID
mylist <- list()
mylist.Bliss <- list()
mylist.HSA <- list()


for (i in 1:nrow(first.two$drug.pairs)) {
  mylist[i] <- first.two$drug.pairs[i,]$blockIDs
}

for (i in 1:nrow(first.two.Bliss$drug.pairs)) {
  mylist.Bliss[i] <- first.two.Bliss$drug.pairs[i,]$blockIDs
}
for (i in 1:nrow(first.two.HSA$drug.pairs)) {
  mylist.HSA[i] <- first.two.HSA$drug.pairs[i,]$blockIDs
}

```

general rename for ZIP.Test
```{r}
colnames(xy.ZIP)[which(names(xy.ZIP) == "Var1")] <- "ConcR"
colnames(xy.ZIP)[which(names(xy.ZIP) == "Var2")] <- "ConcC"
colnames(xy.ZIP)[which(names(xy.ZIP) == "L1")] <- "blockIDs"
colnames(xy.ZIP)[which(names(xy.ZIP) == "value.x")] <- "Response_inhibition"
synergy.type.ZIP <- paste('Synergy', first.two$method, sep = '_')
colnames(xy.ZIP)[which(names(xy.ZIP) == "value.y")] <- synergy.type.ZIP
```
general rename for Bliss.Test
```{r}
colnames(xy.Bliss)[which(names(xy.Bliss) == "Var1")] <- "ConcR"
colnames(xy.Bliss)[which(names(xy.Bliss) == "Var2")] <- "ConcC"
colnames(xy.Bliss)[which(names(xy.Bliss) == "L1")] <- "blockIDs"
colnames(xy.Bliss)[which(names(xy.Bliss) == "value.x")] <- "Response_inhibition"
synergy.type.Bliss <- paste('Synergy', first.two.Bliss$method, sep = '_')
colnames(xy.Bliss)[which(names(xy.Bliss) == "value.y")] <- synergy.type.Bliss
```
general rename for HSA.Test
```{r}
colnames(xy.HSA)[which(names(xy.HSA) == "Var1")] <- "ConcR"
colnames(xy.HSA)[which(names(xy.HSA) == "Var2")] <- "ConcC"
colnames(xy.HSA)[which(names(xy.HSA) == "L1")] <- "blockIDs"
colnames(xy.HSA)[which(names(xy.HSA) == "value.x")] <- "Response_inhibition"
synergy.type.HSA <- paste('Synergy', first.two.HSA$method, sep = '_')
colnames(xy.HSA)[which(names(xy.HSA) == "value.y")] <- synergy.type.HSA
```

now we need to change purely numerical blockIDs with actual blockIDs
```{r}
for (i in 1:nrow(first.two$drug.pairs)) {
  xy.ZIP$blockIDs[xy.ZIP$blockIDs == i] <- mylist[[i]]
}

for (i in 1:nrow(first.two.Bliss$drug.pairs)) {
  xy.Bliss$blockIDs[xy.Bliss$blockIDs == i] <- mylist.Bliss[[i]]
}

for (i in 1:nrow(first.two.HSA$drug.pairs)) {
  xy.HSA$blockIDs[xy.HSA$blockIDs == i] <- mylist.HSA[[i]]
}


```

merging drug names.Test
```{r}
first.two.no.celllines <- merge(xy.ZIP, first.two$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
first.two.no.celllines.Bliss <- merge(xy.Bliss, first.two.Bliss$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
first.two.no.celllines.HSA <- merge(xy.HSA, first.two.HSA$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)


first.two.no.celllines.ZIP.Bliss <- merge(first.two.no.celllines, first.two.no.celllines.Bliss, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
first.two.no.celllines.ZIP.Bliss.HSA <- merge(first.two.no.celllines.ZIP.Bliss, first.two.no.celllines.HSA, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))

```
prep for monstrous loop
```{r}
first.two.response <- today.ZIP$dose.response.mats
first.two.drugs <- today.ZIP$drug.pairs
first.two.scores <- today.ZIP$scores
first.two.method <- today.ZIP$method
first.two <- list()
first.two$dose.response.mats <- first.two.response
first.two$scores <- first.two.scores
first.two$drugs <- first.two.drugs
first.two$method <- first.two.method

first.two.response.Bliss <- today.Bliss$dose.response.mats
first.two.drugs.Bliss <- today.Bliss$drug.pairs
first.two.scores.Bliss <- today.Bliss$scores
first.two.method.Bliss <- today.Bliss$method
first.two.Bliss <- list()
first.two.Bliss$dose.response.mats <- first.two.response.Bliss
first.two.Bliss$scores <- first.two.scores.Bliss
first.two.Bliss$drugs <- first.two.drugs.Bliss
first.two.Bliss$method <- first.two.method.Bliss

first.two.response.L <- today.Loewe$dose.response.mats
first.two.drugs.L <- today.Loewe$drug.pairs
first.two.scores.L <- today.Loewe$scores
first.two.method.L <- today.Loewe$method
first.two.L <- list()
first.two.L$dose.response.mats <- first.two.response.L
first.two.L$scores <- first.two.scores.L
first.two.L$drugs <- first.two.drugs.L
first.two.L$method <- first.two.method.L

first.two.response.H <- today.HSA$dose.response.mats
first.two.drugs.H <- today.HSA$drug.pairs
first.two.scores.H <- today.HSA$scores
first.two.method.H <- today.HSA$method
first.two.H <- list()
first.two.H$dose.response.mats <- first.two.response.H
first.two.H$scores <- first.two.scores.H
first.two.H$drugs <- first.two.drugs.H
first.two.H$method <- first.two.method.H
```
treating NAs (when taking blockID 226).Test
```{r}
#colnames(first.two$dose.response.mats[[2]])
#rownames(first.two$dose.response.mats[[2]])
#dim(first.two$dose.response.mats[[2]])
cols <- colnames(first.two$dose.response.mats[[which(is.na(first.two$scores))]])
rows <- rownames(first.two$dose.response.mats[[which(is.na(first.two$scores))]])
numrows <- dim(first.two$dose.response.mats[[which(is.na(first.two$scores))]])[1]
numcols <- dim(first.two$dose.response.mats[[which(is.na(first.two$scores))]])[2]

first.two$scores[[which(is.na(first.two$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))


```
fuck it, let's do it with loops yo
```{r}
datalist = list()
mylist <- list()
mylist.Bliss <- list()
mylist.HSA <- list()

first.two <- list()
first.two.Bliss <- list()
first.two.HSA <- list()

#populate for ZIP
first.two$dose.response.mats <- reshaped.ZIP$dose.response.mats[c(1:100)]
first.two$drug.pairs <- reshaped.ZIP$drug.pairs[c(1:100),]
first.two$scores <- reshaped.ZIP$scores[c(1:100)]
first.two$method <- reshaped.ZIP$method

#populate for Bliss
first.two.Bliss$dose.response.mats <- reshaped.Bliss$dose.response.mats[c(1:100)]
first.two.Bliss$drug.pairs <- reshaped.Bliss$drug.pairs[c(1:100),]
first.two.Bliss$scores <- reshaped.Bliss$scores[c(1:100)]
first.two.Bliss$method <- reshaped.Bliss$method

#populate for HSA
first.two.HSA$dose.response.mats <- reshaped.HSA$dose.response.mats[c(1:100)]
first.two.HSA$drug.pairs <- reshaped.HSA$drug.pairs[c(1:100),]
first.two.HSA$scores <- reshaped.HSA$scores[c(1:100)]
first.two.HSA$method <- reshaped.HSA$method

for (i in 1:nrow(first.two$drug.pairs)) {
  mylist[i] <- first.two$drug.pairs[i,]$blockIDs
}

for (i in 1:nrow(first.two.Bliss$drug.pairs)) {
  mylist.Bliss[i] <- first.two.Bliss$drug.pairs[i,]$blockIDs
}
for (i in 1:nrow(first.two.HSA$drug.pairs)) {
  mylist.HSA[i] <- first.two.HSA$drug.pairs[i,]$blockIDs
}

  pb <- txtProgressBar(min = 0, max = nrow(first.two$drug.pairs), style = 3)

for (i in 1:nrow(first.two$drug.pairs))
{
      setTxtProgressBar(pb, i)
  a <- list()
  b <- list()
#  c <- list()
  d <- list()
  
  
  a$dose.response.mats <- first.two$dose.response.mats[i]
  a$scores <- first.two$scores[i]
  a$method <- first.two$method
  a$drug.pairs <- first.two$drug.pairs[i,]
  
    if (is.na(a$scores)) {
    cols <- colnames(a$dose.response.mats[[which(is.na(a$scores))]])
    rows <- rownames(a$dose.response.mats[[which(is.na(a$scores))]])
    numrows <- dim(a$dose.response.mats[[which(is.na(a$scores))]])[1]
    numcols <- dim(a$dose.response.mats[[which(is.na(a$scores))]])[2]
    a$scores[[which(is.na(a$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }
  
  b$dose.response.mats <- first.two.Bliss$dose.response.mats[i]
  b$scores <- first.two.Bliss$scores[i]
  b$method <- first.two.Bliss$method
  b$drug.pairs <- first.two.Bliss$drug.pairs[i,]
  
    if (is.na(b$scores)) {
    cols <- colnames(b$dose.response.mats[[which(is.na(b$scores))]])
    rows <- rownames(b$dose.response.mats[[which(is.na(b$scores))]])
    numrows <- dim(b$dose.response.mats[[which(is.na(b$scores))]])[1]
    numcols <- dim(c$dose.response.mats[[which(is.na(b$scores))]])[2]
    b$scores[[which(is.na(b$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }
  
#  c$dose.response.mats <- first.two.L$dose.response.mats[i]
#  c$scores <- first.two.L$scores[i]
#  c$method <- first.two.L$method
#  c$drug.pairs <- first.two.L$drug.pairs[i,]
#  
#    if (is.na(c$scores)) {
#      
#    cols <- colnames(c$dose.response.mats[[which(is.na(c$scores))]])
#    rows <- rownames(c$dose.response.mats[[which(is.na(c$scores))]])
#    numrows <- dim(c$dose.response.mats[[which(is.na(c$scores))]])[1]
#    numcols <- dim(c$dose.response.mats[[which(is.na(c$scores))]])[2]
#    c$scores[[which(is.na(c$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))
#
#  }
  
  d$dose.response.mats <- first.two.HSA$dose.response.mats[i]
  d$scores <- first.two.HSA$scores[i]
  d$method <- first.two.HSA$method
  d$drug.pairs <- first.two.HSA$drug.pairs[i,]
  
    if (is.na(d$scores)) {
    cols <- colnames(d$dose.response.mats[[which(is.na(d$scores))]])
    rows <- rownames(d$dose.response.mats[[which(is.na(d$scores))]])
    numrows <- dim(d$dose.response.mats[[which(is.na(d$scores))]])[1]
    numcols <- dim(d$dose.response.mats[[which(is.na(d$scores))]])[2]
    d$scores[[which(is.na(d$scores))]] <- matrix(,numrows,numcols,dimnames = list(rows, cols))

  }

  x.ZIP <- melt(a$dose.response.mats)
  y.ZIP <- melt(a$scores)
  xy.ZIP <- merge(x.ZIP,y.ZIP, by = c('Var1', 'Var2', 'L1'))
  
  x.Bliss <- melt(b$dose.response.mats)
  y.Bliss <- melt(b$scores)
  xy.Bliss <- merge(x.Bliss,y.Bliss, by = c('Var1', 'Var2', 'L1'))
  
#  x.Loewe <- melt(c$dose.response.mats)
#  y.Loewe <- melt(c$scores)
#  xy.Loewe <- merge(x.Loewe,y.Loewe, by = c('Var1', 'Var2', 'L1'))
  
  x.HSA <- melt(d$dose.response.mats)
  y.HSA <- melt(d$scores)
  xy.HSA <- merge(x.HSA,y.HSA, by = c('Var1', 'Var2', 'L1'))
  
  colnames(xy.ZIP)[which(names(xy.ZIP) == "Var1")] <- "ConcR"
  colnames(xy.ZIP)[which(names(xy.ZIP) == "Var2")] <- "ConcC"
  colnames(xy.ZIP)[which(names(xy.ZIP) == "L1")] <- "blockIDs"
  xy.ZIP$blockIDs <- mylist[[i]]
  colnames(xy.ZIP)[which(names(xy.ZIP) == "value.x")] <- "Response_inhibition"
  synergy.type.ZIP <- paste('Synergy', first.two$method, sep = '_')
  colnames(xy.ZIP)[which(names(xy.ZIP) == "value.y")] <- synergy.type.ZIP
  
  colnames(xy.Bliss)[which(names(xy.Bliss) == "Var1")] <- "ConcR"
  colnames(xy.Bliss)[which(names(xy.Bliss) == "Var2")] <- "ConcC"
  colnames(xy.Bliss)[which(names(xy.Bliss) == "L1")] <- "blockIDs"
  xy.Bliss$blockIDs <- mylist.Bliss[[i]]
  colnames(xy.Bliss)[which(names(xy.Bliss) == "value.x")] <- "Response_inhibition"
  synergy.type.Bliss <- paste('Synergy', first.two.Bliss$method, sep = '_')
  colnames(xy.Bliss)[which(names(xy.Bliss) == "value.y")] <- synergy.type.Bliss
  
#  colnames(xy.Loewe)[which(names(xy.Loewe) == "Var1")] <- "ConcR"
#  colnames(xy.Loewe)[which(names(xy.Loewe) == "Var2")] <- "ConcC"
#  colnames(xy.Loewe)[which(names(xy.Loewe) == "L1")] <- "blockIDs"
#  xy.Loewe$blockIDs <- i
#  colnames(xy.Loewe)[which(names(xy.Loewe) == "value.x")] <- "Response_inhibition"
#  synergy.type.Loewe <- paste('Synergy', first.two.L$method, sep = '_')
#  colnames(xy.Loewe)[which(names(xy.Loewe) == "value.y")] <- synergy.type.Loewe
  
  colnames(xy.HSA)[which(names(xy.HSA) == "Var1")] <- "ConcR"
  colnames(xy.HSA)[which(names(xy.HSA) == "Var2")] <- "ConcC"
  colnames(xy.HSA)[which(names(xy.HSA) == "L1")] <- "blockIDs"
  xy.HSA$blockIDs <- mylist.HSA[[i]]
  colnames(xy.HSA)[which(names(xy.HSA) == "value.x")] <- "Response_inhibition"
  synergy.type.HSA <- paste('Synergy', first.two.HSA$method, sep = '_')
  colnames(xy.HSA)[which(names(xy.HSA) == "value.y")] <- synergy.type.HSA
  
  
  first.two.no.celllines <- merge(xy.ZIP, a$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  first.two.no.celllines.Bliss <- merge(xy.Bliss, b$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
#  first.two.no.celllines.Loewe <- merge(xy.Loewe, c$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  first.two.no.celllines.HSA <- merge(xy.HSA, d$drug.pairs, by.x=c('blockIDs'), by.y = c('blockIDs'), all.x = T)
  
  first.two.no.celllines.ZIP.Bliss <- merge(first.two.no.celllines, first.two.no.celllines.Bliss, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
  first.two.no.celllines.ZIP.Bliss.HSA <- merge(first.two.no.celllines.ZIP.Bliss, first.two.no.celllines.HSA, 
                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
  
#  first.two.no.celllines.ZIP.Bliss.Loewe <- merge(first.two.no.celllines.ZIP.Bliss, first.two.no.celllines.Loewe, 
#                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
#                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
#  first.two.no.celllines.ZIP.Bliss.Loewe.HSA <- merge(first.two.no.celllines.ZIP.Bliss.Loewe, first.two.no.celllines.HSA, 
#                                          by = c('blockIDs','ConcR','ConcC','Response_inhibition',
#                                                 'drug.row', 'drug.col','concRUnit','concCUnit'))
  
  
  
#  datalist[[i]] <- first.two.no.celllines.ZIP.Bliss.Loewe.HSA
  datalist[[i]] <- first.two.no.celllines.ZIP.Bliss.HSA
  
}
    close(pb)
big_data = do.call(rbind, datalist)
```

adding cell lines
```{r}
asdf <- input.sample %>%
  group_by(BlockID) %>%
    transmute(cell_line) %>%
      summarise(cell_line = unique(cell_line))

testing <- big_data
out <- merge(testing, asdf, by.x=c('blockIDs'),by.y=c('BlockID'),all.x = T )
write.csv(out,file = '~/Documents/fimm_files/synergy_calc_august/output_alls.csv')
```



